---
title: "Project 2"
output: html_notebook
author: Isaac Moore
---
```{r}
setwd("~/Google Drive/data_science/general_assembly/Projects/DSI_SM_Project2")
```
Loading packages
```{r}
library("tidyverse")
library("lubridate")
```
##Project 2
###Step 1: Exploring your data.
**Load your data in using Pandas and start to explore. Save all of your early exploration code here and include in your final submission.**
```{r}
data <- tbl_df(read.csv('billboard.csv', stringsAsFactors = F))
head(data)
```

*Viewing the data types in in the dataset.*
```{r}
str(data)
```

*Checking if there are any years ourside of the year 2000.*
```{r}
table(data$year)
```

*Seeing if there are any duplicate artists in the dataset.*  
If there are not duplicate songs - this will tell us which artist had the most songs on the billboard for 2010.
```{r}
head(sort(table(data$artist.inverted), decreasing = T), n = 25)
```

*Checking to see if multiple artists were entered incorrectly or if they had multiple songs on the billboard for 2000.*
```{r}
data[data$artist.inverted == "Jay-Z", ]
```
*Checking to see if multiple artists were entered incorrectly or if they had multiple songs on the billboard for 2000.*
```{r}
data[data$artist.inverted == "Dixie Chicks, The", ]
```

*Chekcing weeks 66 - 76 for any data.*
```{r}
summary(data)
```
Weeks 67 through 76 are all NAs -- 317

*Checking for duplicate songs.*
```{r}
head(sort(table(data$track), decreasing = T))
```

*Checking to see if multiple songs of the same name were by the same artist.*
```{r}
data[data$track == "Where I Wanna Be", ]
```

*Checking to see what is the most popular genre of music.*
```{r}
sort(table(data$genre), decreasing = T)
```

*Shortest song & longest song*
```{r}
data$time <- as.character(data$time)
min(data$time)
max(data$time)
```


*Shortest song*
```{r}
data[data$time == "2:36", ]
```

*Longest song*
```{r}
data[data$time == "7:50",]
```

*Converting the time of the track to seconds.*
```{r}
convert_time <- function(x){
        tmp <- strsplit(x, ":")
        return((as.integer(tmp[[1]][1]) * 60) + as.integer(tmp[[1]][2]))
}
data$time_seconds <- lapply(data$time, convert_time)
```

*Finding the most popular track for 2010 - in terms of amount of time on the billboard.*  
First I will tidy the data, using the Hadley Wickam's dity data principles.
```{r}
data2 <- data %>% 
  tidyr::gather(week, rank, x1st.week:x76th.week, na.rm = TRUE)
```
Since each observation is a week on the dataset, all I have to do now is create a srted table of the artists.
```{r}
popular_tracks <- sort(table(data2$track), decreasing = T)
head(popular_tracks, n = 15)
```

*Finding the average amount of time a songs was on the billboard.*
```{r}
popular_tracks2 <- c()
for (i in popular_tracks){
        # print(i)
  popular_tracks2 <- c(popular_tracks2, i)
}
summary(popular_tracks2)
```

*#1 tracks for 2010*
```{r}
top_tracks <- data2 %>% filter(rank == 1) 
top_tracks
```
```{r}
top_tracks %>% count(artist.inverted) %>% arrange(desc(n))
top_tracks %>% count(track) %>% arrange(desc(n))
```

###Step 2: Clean your data.
Do some rudimentary cleaning. Rename any columns that are poorly named, shorten any strings that may be too long, fill in any missing values. Explain your rationale for the way you choose to "impute" the missing data.
```{r}
colnames(data2) <- c("year", "artists", "track", "time", "genre", "date_enter", "date_peak", "time_secs", "week", "rank")
colnames(data2)
```

```{r}
data2$time_secs <- as.numeric(data2$time_secs)
summary(data2$time_secs)
```

```{r}
class(data2$date_enter)
class(data2$date_peak)
```
```{r}
data2$date_enter <- ymd(data2$date_enter)
data2$date_peak <- ymd(data2$date_peak)
```

```{r}
class(data2$date_enter)
class(data2$date_peak)
```

```{r}
glimpse(data2)
```

### Step 3: Visualize your data.
*Using a plotting utility of your choice (Tableau or python modules or both), create visualizations that will provide context to your data. There is no minimum or maximum number of graphs you should generate, but there should be a clear and consistent story being told. Give insights to the distribution, statistics, and relationships of the data.*

```{r}
# Visualize most popular genre
ggplot(data2, aes(genre)) + geom_histogram(stat = "count") + ggtitle('Number of Songs per Genre on the \n Billboard Top 100 in 2010')
```

```{r}
# Correlation between song length and rank
time_on_bb <- data2 %>% select(track, time_secs, rank) %>% 
        group_by(track) %>% 
                mutate(time_on_bbs = n()) %>% 
                        arrange(desc(time_on_bbs))

ggplot(time_on_bb, aes(time_secs, time_on_bbs)) + geom_point() + geom_smooth(method = lm) + ggtitle('Correlation between song length (seconds)\n and time (weeks) on the Billboard Top 100') + ylab("Weeks on Billboard Top 100") + xlab("Length of track (seconds)")
```

```{r}
# Genres and mean length of song
avg_song_genre <- data2 %>% group_by(genre) %>% summarise(avg_length = mean(time_secs))
ggplot(avg_song_genre, aes(x = genre, y = avg_length)) + geom_bar(stat = "identity") + labs(title='Average song length per Genre', ylabel='Average length of song (seconds)', xlabel=' Genre')
```

```{r}
# Independent Women
ind_wom <- filter(data2, track == "Independent Women Part I") 
ind_wom$week <- readr::parse_number(ind_wom$week)
ggplot(ind_wom, aes(x = week, y = rank)) + geom_line() + labs(title='Destiny\'s Child - Independent Women Part I\n time on the Billboard Top 100', xlabel='Weeks', ylabel='Rank')
```


