---
title: "Project 2"
output: html_notebook
author: Isaac Moore
---
```{r}
setwd("~/Google Drive/data_science/general_assembly/Projects/DSI_SM_Project2")
```
Loading packages
```{r}
library("tidyverse")
library("lubridate")
```
##Project 2
###Step 1: Exploring your data.
**Load your data in using Pandas and start to explore. Save all of your early exploration code here and include in your final submission.**
```{r}
data <- tbl_df(read.csv('billboard.csv'))
head(data)
```

*Viewing the data types in in the dataset.*
```{r}
str(data)
```

*Checking if there are any years ourside of the year 2000.*
```{r}
table(data$year)
```

*Seeing if there are any duplicate artists in the dataset.*  
If there are not duplicate songs - this will tell us which artist had the most songs on the billboard for 2010.
```{r}
head(sort(table(data$artist.inverted), decreasing = T), n = 25)
```

*Checking to see if multiple artists were entered incorrectly or if they had multiple songs on the billboard for 2000.*
```{r}
data[data$artist.inverted == "Jay-Z", ]
```
*Checking to see if multiple artists were entered incorrectly or if they had multiple songs on the billboard for 2000.*
```{r}
data[data$artist.inverted == "Dixie Chicks, The", ]
```

*Chekcing weeks 66 - 76 for any data.*
```{r}
summary(data)
```
Weeks 67 through 76 are all NAs -- 317

*Checking for duplicate songs.*
```{r}
head(sort(table(data$track), decreasing = T))
```

*Checking to see if multiple songs of the same name were by the same artist.*
```{r}
data[data$track == "Where I Wanna Be", ]
```

*Checking to see what is the most popular genre of music.*
```{r}
sort(table(data$genre), decreasing = T)
```

*Shortest song & longest song*
```{r}
data$time <- as.character(data$time)
min(data$time)
max(data$time)
```


*Shortest song*
```{r}
data[data$time == "2:36", ]
```

*Longest song*
```{r}
data[data$time == "7:50",]
```

*Converting the time of the track to seconds.*
```{r}
convert_time <- function(x){
        tmp <- strsplit(x, ":")
        return((as.integer(tmp[[1]][1]) * 60) + as.integer(tmp[[1]][2]))
}
data$time_seconds <- lapply(data$time, convert_time)
```

*Finding the most popular track for 2010 - in terms of amount of time on the billboard.*
```{r}

```

